\documentclass{article}
\begin{document}

\section{Data-Intensive Categories}

\begin{enumerate}
    \item \textbf{Machine Learning (ML) and Deep Learning}
    \begin{itemize}
        \item \textbf{Application}: Used for tasks such as image recognition, natural language processing, recommendation systems, and autonomous vehicles.
        \item \textbf{Key Concepts}: Neural networks, backpropagation, model training, inference, GPU/TPU acceleration.
        \item \textbf{Resources}:
        \begin{itemize}
            \item \textit{Deep Learning} by Ian Goodfellow and Yoshua Bengio.
            \item \textit{Hands-On Machine Learning with Scikit-Learn, Keras, and TensorFlow} by Aurélien Géron.
            \item \textit{Fast.ai} (Practical deep learning online course).
        \end{itemize}
    \end{itemize}
    
    \item \textbf{Big Data Analytics}
    \begin{itemize}
        \item \textbf{Application}: Processing massive datasets for insights, often used in social media platforms, e-commerce, and healthcare.
        \item \textbf{Key Concepts}: Parallel processing, data lakes, distributed computing.
        \item \textbf{Technologies}: Apache Hadoop, Apache Spark.
        \item \textbf{Resources}:
        \begin{itemize}
            \item \textit{Hadoop: The Definitive Guide} by Tom White.
            \item \textit{The Data Engineering Cookbook} by Andreas Kretz.
        \end{itemize}
    \end{itemize}

    \item \textbf{Blockchain and Cryptocurrency}
    \begin{itemize}
        \item \textbf{Application}: Decentralized systems used in cryptocurrencies like Bitcoin and Ethereum.
        \item \textbf{Key Concepts}: Proof-of-work, proof-of-stake, cryptographic hashing.
        \item \textbf{Resources}:
        \begin{itemize}
            \item \textit{Mastering Bitcoin} by Andreas M. Antonopoulos.
            \item \textit{Blockchain Basics} by Daniel Drescher.
        \end{itemize}
    \end{itemize}

    \item \textbf{Cloud Computing}
    \begin{itemize}
        \item \textbf{Application}: Scalable on-demand resources for storage, computation, and networking.
        \item \textbf{Key Concepts}: IaaS, PaaS, serverless architecture.
        \item \textbf{Technologies}: AWS, Microsoft Azure, Google Cloud.
        \item \textbf{Resources}:
        \begin{itemize}
            \item \textit{Cloud Computing: Concepts, Technology \& Architecture} by Thomas Erl.
            \item AWS training resources and free-tier experimentation.
        \end{itemize}
    \end{itemize}

    \item \textbf{Distributed Databases}
    \begin{itemize}
        \item \textbf{Application}: Databases designed for large datasets with fault tolerance and scalability.
        \item \textbf{Key Concepts}: Sharding, replication, consistency models.
        \item \textbf{Technologies}: Cassandra, MongoDB, DynamoDB.
        \item \textbf{Resources}:
        \begin{itemize}
            \item \textit{Designing Data-Intensive Applications} by Martin Kleppmann.
            \item \textit{Cassandra: The Definitive Guide} by Eben Hewitt.
        \end{itemize}
    \end{itemize}

    \item \textbf{Edge Computing}
    \begin{itemize}
        \item \textbf{Application}: Real-time processing at the network's edge for IoT devices and smart cities.
        \item \textbf{Key Concepts}: Low-latency networks, 5G, fog computing.
        \item \textbf{Resources}:
        \begin{itemize}
            \item \textit{Architecting the Internet of Things} by Dieter Uckelmann et al.
            \item IBM and AWS IoT Edge services.
        \end{itemize}
    \end{itemize}

    \item \textbf{AI in Cloud Computing}
    \begin{itemize}
        \item \textbf{Application}: Cloud services to build, train, and deploy machine learning models.
        \item \textbf{Technologies}: AWS SageMaker, Google AI Platform.
        \item \textbf{Resources}:
        \begin{itemize}
            \item \textit{Architecting AI in the Cloud} by Lee Atchison.
            \item Online AWS and Google Cloud AI courses.
        \end{itemize}
    \end{itemize}

    \item \textbf{High-Performance Computing (HPC)}
    \begin{itemize}
        \item \textbf{Application}: Used in simulations, scientific computing, climate modeling, and physics simulations.
        \item \textbf{Key Concepts}: Distributed memory, parallel computing, task scheduling.
        \item \textbf{Resources}:
        \begin{itemize}
            \item \textit{Introduction to High-Performance Computing for Scientists and Engineers} by Georg Hager.
            \item \textit{Parallel Programming in MPI and OpenMP} by Peter Pacheco.
        \end{itemize}
    \end{itemize}

    \item \textbf{Data Streaming and Real-Time Analytics}
    \begin{itemize}
        \item \textbf{Application}: Real-time data processing, commonly used in financial markets and live monitoring systems.
        \item \textbf{Key Concepts}: Stream processing, event-driven architecture.
        \item \textbf{Technologies}: Apache Kafka, Flink.
        \item \textbf{Resources}:
        \begin{itemize}
            \item \textit{Kafka: The Definitive Guide} by Neha Narkhede.
            \item \textit{Streaming Systems} by Tyler Akidau et al.
        \end{itemize}
    \end{itemize}

    \item \textbf{Autonomous Systems}
    \begin{itemize}
        \item \textbf{Application}: Self-driving cars, drones, robotics.
        \item \textbf{Key Concepts}: Sensor fusion, reinforcement learning, computer vision.
        \item \textbf{Resources}:
        \begin{itemize}
            \item \textit{Robotics: Modelling, Planning and Control} by Bruno Siciliano.
            \item \textit{Self-Driving Cars: The Future is Here} by Steven Shladover.
        \end{itemize}
    \end{itemize}
\end{enumerate}

\end{document}
